Comparison of MCMC algorithms in Stochastic Volatility Models

### 2023-08-29 writeups/proposal_presentation/proposal_presentation_sv.qmd
background

## Background
- Hello

@@

- 3 key components of research
    + Kim, Sherphard and Chib (1998) SV model
    + Hamiltonian Monte Carlo
    + Simulation Based Calibration to validate MCMC computation

@@

$$
\begin{aligned}
y_t =& \space \beta exp(h_t/2) \epsilon_t \\
h_{t+1} =& \space \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t  \\
h_1 \sim& \space normal\left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right) \\
\end{aligned}
$$

@@

- $h_t$ is log volatility where $h_1$ is a draw from a stationary distribution and the state equation $h_{t+1}$ follows a stationary process governed by the autoregressive parameter $\phi$ such that $|\phi|<1$.

@@

## Next steps
- Consider different model parameterisation
    + Performance of a sampler may be sensititive to the shape of the posterior density

- Consider KSC's correction using importance sampling weights (MH correction step)
    + Reweights samples from approximate distribution such that reweighted draws come from the correct distribution

>>>>

## Background
- Variety of Bayesian estimation strategies for different types of Stochastic Volatility models using classical Markov Chain Monte Carlo (MCMC) algorithms

- Since then, more advanced algorithms such as Hamiltoninan Monte Carlo has enabled the estimation of richer and more complicated models

- However, as the developments of these tools enable the estimation of more complicated models so do the risk of mistakes and the need to validate the output of our algorithms

- How do we ensure the validity of inferences from our algorithms across different estimation strategies?

<!-- What is the fundamental difference between convergence diagnostics, bayes factors with SBC? Check the fit of the model?-->

@@

- 3 key components
    + Kim, Sherphard and Chib (1998) SV model
    + Hamiltonian Monte Carlo
    + Simulation Based Calibration to validate MCMC computation

@@

$$
\begin{aligned}
y_t =& \space \beta exp(h_t/2) \epsilon_t \\
h_{t+1} =& \space \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t  \\
h_1 \sim& \space normal\left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right) \\
\epsilon_t \sim &N(0,1) \space\space\space \eta_t \sim N(0,1)
\end{aligned}
$$

@@

- $h_t$ is log volatility where $h_1$ is a draw from a stationary distribution and the state equation $h_{t+1}$ is a stationary process governed by the autoregressive parameter $\phi$ where $|\phi|<1$.

@@

## Next steps

- Increasing number of posterior draws for KSC model
    + MCMC may not have converged onto target distribution

- Different model parameterisation
    + Performance of a sampler may be sensititive to the shape of the posterior density

- KSC's correction using importance sampling weights (MH correction step)
    + Reweights samples from approximate distribution such that reweighted draws come from the correct distribution
